---
title: "FA7 Part 1"
author: "Angelo M. Navarro"
date: "2026-10-08"
output: html_document
---
## Importing & Extracting Dataset
```{r}
library(readxl)
Elevator_Waiting_Times <- read_excel("C:/Users/Angelo/OneDrive - Far Eastern University/Prob/Elevator Waiting Times.xlsx")
View(Elevator_Waiting_Times)

wait <- Elevator_Waiting_Times$`Waiting Time`
```

# Key Parameters

## Mean
E(X) = 1 / λ
```{r}
mean(wait)
```
mean(wait) = 1 / λ
λ = 1 / mean(wait)

```{r}
lambda_wait <- 1 / mean(wait)
```

*On average, an FC Building elevator arrives once every 4.0667 minutes.*

## Additional Statistical Measures
```{r}
min(wait)
max(wait)
```


## Probability Density Function

```{r}
PDF <- function(x, lambda) {
  result <- (lambda * exp(-(lambda) * x))
  return(result)
  }
```

Let's test:

```{r}
PDF(wait, lambda_wait)
dexp(wait, lambda_wait)
```

## Cumulative Density Function

```{r}
CDF <- function(x, lambda) {
  result <- (1 - exp(-(lambda) * x))
  return(result)
  }
```

```{r}
CDF(wait, lambda_wait)
pexp(wait, lambda_wait)
```

## Probabilities
```{r}
fd_wait <- sort(unique(wait), decreasing = TRUE)
fd_tab <- table(dexp(wait, lambda_wait))
data.frame(fd_wait, fd_tab)
```

## Visualizing the Data
```{r}
curve(dexp(x, lambda_wait), from = min(wait), to = max(wait), xlab = " t " , ylab = "f(t)", main = "Elevator Waiting Time")
```
```{r}
ind <- c(1:10)
data.frame(ind, dexp(ind, lambda_wait))
```

```{r}
hist(dexp(wait, lambda_wait), xlab = "Probability", ylab = "Minutes", ylim = c(1, 10), main = "Probabilities of Wait Times")
```

# Interpretations
The exponential distribution allows us to determine the *likelihood that the elevator arrives within a specific number of minutes*. Using the CDF, we can calculate the probability that a student waits less than or equal to a given number of seconds. The expected waiting time, calculated as 1/λ represents the average amount of time a student can expect to wait for the elevator. These results can help students anticipate waiting times and provide insight into the efficiency of the elevator service.

## Implications
How likely is a wait time of exactly 1 minute?
```{r}
dexp(1, lambda_wait)
```

**19.23% likelihood.**

How likely is a wait time of 3 minutes or less?
```{r}
pexp(3, lambda_wait)
```
**52.18% likelihood.**

How likely is a wait time of exactly 10 minutes?
```{r}
dexp(10, lambda_wait)
```
**2.10% Likelihood.**